Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-26 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
edgb appendix and conclusions

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
Some of these gravitational-wave events have sufficiently high
signal-to-noise ratio in the merger, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown, which is
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we  address this question by using an inspiral-merger-ringdown waveform
model in the effective-one-body formalism, augmented with a parametrization of
the ringdown based on an expansion in the final black hole's spin.
%
We give a prescription on how to include in this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis. In particular, we focus on theories that modify
general relativity by higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet, dynamical Chern-Simons, cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against the two loudest
gravitational-wave ringdown signals observed so far, notably, GW150914 and GW200129.
%
We find that while the Einstein-dilaton-Gauss-Bonnet theory cannot be
constrained with these events, we can place upper bounds on the
fundamental length-scale of cubic- ($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order
($\ell_{\rm qEFT} \leqslant 51.3$~km) effective-field-theories of general
relativity, and of dynamical Chern-Simons gravity ($\ell_{\rm dCS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently unconstrained by
inspiral-only analyses which, however, can be constrained by merger-ringdown
studies with current gravitational-wave data.
%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

The large number of GW observations has allowed us to infer relevant astrophysical~\cite{LIGOScientific:2021aug} and cosmological~\cite{LIGOScientific:2021psn} information on the compact-object population in our local Universe, and also to probe General Relativity (GR) in the high-velocity, dynamical and strong-field regime of gravity~\cite{LIGOScientific:2021sio}. The latter complement tests of GR in the low-velocity, quasi-static or linear regimes available with Solar-System experiments~\cite{Will:2014kxa}, binary-pulsar~\cite{Wex:2014nva,Kramer:2021jcw} and galactic-center~\cite{GRAVITY:2018ofz,Do:2019txf,EventHorizonTelescope:2019ths} observations,
and cosmological measurements~\cite{Clifton:2011jh}.

@@

BHs in EdGB gravity have scalar hair, to which we can associate a (monopole) scalar charge $Q$,
related to the asymptotic $r^{-1}$ fall-off of the scalar field, where $r$ is the distance from the BH.
%
This charge is not an independent parameter, and depends on the BH's mass and spin,
thus being a ``secondary hair''~\cite{Coleman:1991ku,Kanti:1995vq,Herdeiro:2015waa}.
%
Since the scalar field is sourced by a curvature scalar, $Q$ is larger (smaller),
the smaller (larger) the BH mass is.

These properties are not mere theoretical curiosities; they have important observational consequences.
%
First, the presence of the scalar charge implies that when in binaries, BHs can source
scalar-dipole radiation (see, e.g., Refs.~\cite{Yagi:2011xp,Julie:2019sab,Shiralilou:2020gah,Shiralilou:2021mfl,Julie:2022huo})
which affects the GW phase at $-1$PN order (relative to the dominant quadrupolar GR
contribution), with magnitude proportional to the difference between the
charges of binary components. This makes EdGB gravity testable with present GW
observations of compact binaries where at least one component is a BH.
%
More specifically, Ref.~\cite{Lyu:2022gdr} placed the bound
%
$\ell_{\rm EdGB} \leqslant 1.33$~km
%
using the NSBH binaries GW200105 and
GW200115~\cite{LIGOScientific:2021qlt} while Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, obtained
%
$\ell_{\rm EdGB} \leqslant 1.7$~km,
%
by stacking the posteriors of $\ell_{\rm EdGB}$ from a selection of BBHs
from the GWTC-1 and GWTC-2 catalogs~\cite{LIGOScientific:2018mvr,LIGOScientific:2020ibl}\footnote{\hs{These bounds
are, strictly speaking, valid only when the scalar field $\varphi$ is small (i.e., $\varphi \ll 1$), and
take into account only the leading-order scalar field interaction arising from the original dilatonic coupling, i.e.,
$f(\varphi) \approx \varphi$ in Eq.~\eqref{eq:action_sgb}.
This results in what is often referred to as shift-symmetric scalar-Gauss-Bonnet theory.
In this theory, NSs do not have scalar monopole charge~\cite{Yagi:2015oca}, while BHs do~\cite{Yunes:2011we,Sotiriou:2013qea,Sotiriou:2014pfa}.}}.
Secondly, the scalar field influences the response of BHs to linearized perturbations and thus
affects the BH's QNM spectra.
%
The coupling between scalar field and the Gauss-Bonnet invariant, which
result in a coupling between scalar perturbations and gravitational
perturbations of polar parity can, for example break the equivalence
between the QNM spectra of polar and axial gravitational perturbations~\cite{Glampedakis:2017rar,Lenzi:2021njy}
(i.e., the isospectrality~\cite{Chandrasekhar:1985kt}) for Schwarzschild BHs in GR.

@@

This theory is described by the action~\cite{Jackiw:2003pm,Alexander:2009tp},
%
\begin{align} \label{eq:action_dcs}
    S_{\rm dCS} = \frac{1}{16 \pi}
    \int \dV
    \left[
    R - \tfrac{1}{2}(\partial \vartheta)^2
    + \tfrac{1}{4} \ell^{2}_{\rm dCS} \, \vartheta \, {}^{*}RR
    \right],
    \nonumber \\
\end{align}
%
where $\vartheta$ is a pseudoscalar field which couples to the Pontryagin
density
${}^{*}RR = {}^{*}R_{\mu\nu}{}^{\rho\sigma} R^{\nu\mu}{}_{\rho\sigma}$,
%
where ${}^{*}R^{\mu\nu\rho\sigma}$ is the dual of the Riemann tensor
defined as
%
${}^{*}R^{\mu\nu\rho\sigma} =
\epsilon^{\mu\nu}{}_{\gamma\delta}
R^{\gamma\delta\rho\sigma} / 2$,
%
and $\epsilon_{\mu\nu\gamma\delta}$ is the Levi-Civita tensor.
%
As in scalar-Gauss-Bonnet theories, $\int \dV {}^{*}RR$ by itself is a boundary
term~\cite{Jackiw:2003pm} that does not contribute to the field equations in
four-dimensions.
%
The Pontryagin density can, however, modify the field equation when coupled to $\vartheta$;
the strength of the coupling set by $\ell_{\rm dCS}$, with dimensions of length.
%
% However, a coupling with $\vartheta$ can modify the field equations.
% %
% The coupling strength is set by the coupling constant $\ell_{\rm dCS}$, with
% dimensions of length.

@@

\begin{table*}[t]
% \begin{tabular}{c | c c c c c c c c c c}
% \hline
% \hline
% Theory      & $\omega^{(0)}_{220}$ & $\tau^{(0)}_{220}$ & $\omega^{(1)}_{220}$ & $\tau^{(1)}_{220}$ & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
% \hline
% GR          & 0.3737               & 11.2407            &  0.1258              & 0.2522             & --  & --                          & --                        & --                          & --                        & \cite{Maselli:2019mjd} \\
% \hline
% EdGB        & --                   & --                 &  --                  & --                 & 4   & 0.0107                      & 0.0044                    & $-$0.2480                   & $-$1.101                  & \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
% dCS         & --                   & --                 &  --                  & --                 & 4   & 3.1964                      & 6.3619                    & 41.199                      & 794.66                    & \cite{Wagle:2021tam}  \\
% cubic EFT   & --                   & --                 &  --                  & --                 & 4   & $-$0.5813                   & 2.6469                    & $-$3.8620                   & 265.12                    & \cite{Cano:2021myl} \\
% quartic EFT & --                   & --                 &  --                  & --                 & 6   & $-$0.2114                   & $-$0.6070                 & $-$1.5263                   & 171.35                    & \cite{Cano:2021myl}  \\
% \hline
% \hline
% \end{tabular}
%
\begin{tabular}{c | c }
\hline
\hline
Theory                  & GR~\cite{Maselli:2019mjd} \\
\hline
$\omega^{(0)}_{220}$    & 0.3737                    \\
$\tau^{(0)}_{220}$      & 11.2407                   \\
$\omega^{(1)}_{220}$    & 0.1258                    \\
$\tau^{(1)}_{220}$      & 0.2522                    \\
\hline
\hline
\end{tabular}
\quad
\begin{tabular}{c | c  c  c  c}
\hline
\hline
Theory                        & EdGB ($p=4$)~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} & dCS ($p=4$)~\cite{Wagle:2021tam} & cubic EFT ($p=4$)~\cite{Cano:2021myl} & quartic EFT ($p=6$)~\cite{Cano:2021myl} \\
\hline
$\delta\omega^{(0)}_{220}$    & 0.0107                                               & 3.1964                   & $-$0.5813                     & $-$0.2114                       \\
$\delta\tau^{(0)}_{220}$      & 0.0044                                               & 6.3619                   & $-$0.2114                     & $-$0.6070                       \\
$\delta\omega^{(1)}_{220}$    & $-$0.2480                                            & 41.199                   & $-$3.8620                     & $-$1.5263                       \\
$\delta\tau^{(1)}_{220}$      & $-$1.1010                                            & 794.66                   & 265.12                        & 171.35                          \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
%
We also included for comparison the GR coefficients, up to $j=1$, obtained in Ref.~\cite{Maselli:2019mjd}.
%
The remaining GR coefficients for $1< j \leqslant 4$, for which their non-GR counterpart cannot be determined as of yet for
the theories under consideration, can be found in Table~I of Ref.~\cite{Maselli:2019mjd}.
}
\label{tab:ref_theories_qnms}
\end{table*}

@@

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that it should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to \hs{each theory's fundamental length-scale} $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
the Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final results using
$\varepsilon = 1/2$.
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %
We say that a bound has been placed on $\ell_{\rm th}$, if most of the PDF $P(\ell_{\rm th} | d)$ support is in the interval
$[0, \, \Lambda_{\rm EFT}(1/2, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell_{\rm th} | d)$, namely
%
\begin{equation}
P(\ell_{\rm th} \leqslant \ell^{\rm max}_{\rm th} | d) = \int_{0}^{\,\ell^{\rm max}_{\rm th}} \dd \ell'\, P(\ell' | d).
\label{eq:def_cdf}
\end{equation}
%
For instance, we require that for a bound at 90\% credible level to be placed on $\ell_{\rm th}$ that
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell_{\rm th} \leqslant \Lambda_{\rm EFT} | d) \leqslant 0.9,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
where we let $\ell^{\rm max}_{\rm th} = \Lambda_{\rm EFT}$ in Eq.~\eqref{eq:def_cdf}, and
likewise for other credibility percentiles.

@@

We start with EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the marginalized PDFs of the
coupling constant $\ell_{\rm GB}$, for GW150914 (top panel) and GW200129
(middle panel), with and without the spin corrections to the $(2,2,0)$ QNM.
%
The bottom panel shows the joint posterior obtained by combining both events.
%
We see, for both events, the $N_{\rm max} = 0$ posteriors are
characterized by a peak away from zero.
%
This does not mean that we are inferring a deviation from GR.
%
We recall that the deviations from GR in the ParSpec framework are controlled
by the dimensionless parameter $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm EdGB} = \left(\frac{c^2 \ell_{\rm EdGB}}{ G M_{f}}\right)^4\,.
    \label{eq:def_gamma_gb}
\end{equation}
%
As shown in Fig.~\ref{fig:gamma_gb}, $\gamma_{\rm EdGB}$ does indeed have a posterior
distribution with largest support at zero, indicating consistency between
the underlying signal and GR.
%
We also observe that the inclusion of the spin corrections (i.e., the curves with $N_{\rm max} = 1$)
displaces the posteriors distributions towards smaller values of $\ell_{\rm EdGB}$ (see Fig.~\ref{fig:edgb_exec_sum}),
and larger values of $\gamma_{\rm EdGB}$ (see Fig.~\ref{fig:gamma_gb}).

@@

As we have emphasized in Sec.~\ref{sec:remarks}, we must first check whether the
``EFT''~\eqref{eq:eft_bound} and ``ParSpec''~\eqref{eq:parspec_bound} bounds
are satisfied, before drawing any conclusions on the allowed values for
$\ell_{\rm EdGB}$ from our parameter estimations.
%
We check the validity of the EFT bound in Fig.~\ref{fig:edgb_cdf}. In the top (bottom) panel we show
the CDF of the $\ell_{\rm EdGB}$ posteriors for GW150914 (GW200129), obtained
by evaluating the integral~\eqref{eq:def_cdf} with $\ell^{\rm max}_{\rm th} = \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or
the remnant's mass ($\mathfrak{m} = M_{\rm f}$, solid lines), while varying $\varepsilon$ between 0 and 1.
%
For GW150914, we see that for the $N_{\rm max} = 0$ curves, that the CDF never goes
past $0.2$, regardless of the mass scale $\mathfrak{m}$ used and even at $\varepsilon =
1$, at which the EFT description of the theory would not be valid anyway.
%
This shows that that the ``EFT bound'' given by Eq.~\eqref{eq:eft_bound} is
never met at a significant credibility and thus that we cannot place a bound on
$\ell_{\rm EdGB}$.
%
The situation is similar for GW200129 with $N_{\rm max} = 0$ and does not
change for either event when we add spin corrections to the EdGB QNM.
%
For the case with $N_{\rm max} = 1$, we find that the ``EFT bound'' is satisfied
only for $\varepsilon \approx 0.8$ and $\approx1$ for GW150914 and GW200129, respectively.
However, we set the maximum value of $\varepsilon$ to be 1/2, thus,
taken together we are led to \emph{conclude that we cannot constrain EdGB gravity with our
present model}.
%
We summarize our findings in Table~\ref{tab:summary_edgb}.

@@

For dCS gravity, we follow Ref.~\cite{Wagle:2021tam}, which numerically calculated the QNMs, $\sigma_{\rm dCS}$,
of slowly-rotating BHs, and found that for the axial gravitational modes the damping time increases, as we
increase $\ell_{\rm dCS}$, \hs{at constant BH spin.}
%
Hence, according to the prescription of Sec.~\ref{sec:theory_specific_qnm},
this is the branch of QNMs we choose to work with.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
Some of these gravitational-wave events have sufficiently high
signal-to-noise ratio in the merger, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown, which is
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we  address this question by using an inspiral-merger-ringdown waveform
model in the effective-one-body formalism, augmented with a parametrization of
the ringdown based on an expansion in the final black hole's spin.
%
We give a prescription on how to include in this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis. In particular, we focus on theories that modify
general relativity by higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet and dynamical Chern-Simons theories, and cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against the two loudest
gravitational-wave ringdown signals observed so far, notably, GW150914 and GW200129.
%
We find that while the Einstein-dilaton-Gauss-Bonnet theory cannot be
constrained with these events, we can place upper bounds on the
fundamental length-scale of cubic- ($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order
($\ell_{\rm qEFT} \leqslant 51.3$~km) effective-field-theories of general
relativity, and of dynamical Chern-Simons gravity ($\ell_{\rm dCS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently unconstrained by
inspiral-only analyses which, however, can be constrained by merger-ringdown
studies with current gravitational-wave data.
%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

The large number of GW observations has allowed us to infer relevant astrophysical~\cite{LIGOScientific:2021aug} and cosmological~\cite{LIGOScientific:2021psn} information on the compact-object population in our local Universe, and also to probe General Relativity (GR) in the high-velocity, dynamical and strong-field regime of gravity~\cite{Yunes:2016jcc,LIGOScientific:2021sio}. The latter complement tests of GR in the low-velocity, quasi-static or linear regimes available with Solar-System experiments~\cite{Will:2014kxa}, binary-pulsar~\cite{Wex:2014nva,Kramer:2021jcw} and galactic-center~\cite{GRAVITY:2018ofz,Do:2019txf,EventHorizonTelescope:2019ths} observations,
and cosmological measurements~\cite{Clifton:2011jh}.

@@

BHs in EdGB gravity have scalar hair, to which we can associate a monopole scalar charge,
related to the asymptotic $r^{-1}$ fall-off of the scalar field, where $r$ is the distance from the BH.
%
This charge is not an independent parameter, and depends on the BH's mass and spin,
thus being a ``secondary hair''~\cite{Coleman:1991ku,Kanti:1995vq,Herdeiro:2015waa}.
%
Since the scalar field is sourced by a curvature scalar, the scalar charge is larger (smaller),
the smaller (larger) the BH mass is.

These properties are not mere theoretical curiosities; they have important observational consequences.
%
First, the presence of the scalar charge implies that when in binaries, BHs can source
scalar-dipole radiation (see, e.g., Refs.~\cite{Yagi:2011xp,Julie:2019sab,Shiralilou:2020gah,Shiralilou:2021mfl,Julie:2022huo})
which affects the GW phase at $-1$PN order (relative to the dominant quadrupolar GR
contribution), with magnitude proportional to the difference between the
charges of binary components. This makes EdGB gravity testable with present GW
observations of compact binaries where at least one component is a BH.
%
More specifically, Ref.~\cite{Lyu:2022gdr} placed the bound
%
% $\ell_{\rm EdGB} \leqslant 1.33$~km
\hs{$\ell_{\rm EdGB} \leqslant 7.1$~km}
%
using the NSBH binaries GW200105 and
GW200115~\cite{LIGOScientific:2021qlt} while Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, obtained
%
% $\ell_{\rm EdGB} \leqslant 1.7$~km,
\hs{$\ell_{\rm EdGB} \leqslant 9.1$~km},
%
by stacking the posteriors of $\ell_{\rm EdGB}$ from a selection of BBHs
from the GWTC-1 and GWTC-2 catalogs~\cite{LIGOScientific:2018mvr,LIGOScientific:2020ibl}\footnote{\hs{These bounds
are, strictly speaking, valid only when the scalar field $\varphi$ is small (i.e., $\varphi \ll 1$), and
take into account only the leading-order scalar field interaction arising from the original dilatonic coupling, i.e.,
$f(\varphi) \approx \varphi$ in Eq.~\eqref{eq:action_sgb}.
This results in what is often referred to as shift-symmetric scalar-Gauss-Bonnet theory.
In this theory, NSs do not have scalar monopole charge~\cite{Yagi:2015oca}, while BHs do~\cite{Yunes:2011we,Sotiriou:2013qea,Sotiriou:2014pfa}. Finally, we take note constant $\alpha_{\rm EdGB}$ used in Refs.~\cite{Yagi:2011xp,Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} is related to $\ell_{\rm EdGB}$ as $\ell_{\rm EdGB} = 4 \pi^{1/4} |\alpha_{\rm EdGB}|^{1/2}$.
\hscomm{I had to translate between notations and rewrote the bounds quoted for $\alpha_{\rm EdGB}$ in terms of our $\ell_{\rm EdGB}$.}
}}.
Secondly, the scalar field influences the response of BHs to linearized perturbations and thus
affects the BH's QNM spectra.
%
The coupling between scalar field and the Gauss-Bonnet invariant, which
result in a coupling between scalar perturbations and gravitational
perturbations of polar parity can, for example break the equivalence
between the QNM spectra of polar and axial gravitational perturbations~\cite{Glampedakis:2017rar,Lenzi:2021njy}
(i.e., the isospectrality~\cite{Chandrasekhar:1985kt}) for Schwarzschild BHs in GR.

@@

This theory is described by the action~\cite{Jackiw:2003pm,Alexander:2009tp},
%
\begin{align} \label{eq:action_dcs}
    S_{\rm dCS} =
    \int \dV
    \left[
    \hs{\tfrac{1}{16\pi}} R - \tfrac{1}{2}(\partial \vartheta)^2
    + \tfrac{1}{4} \ell^{2}_{\rm dCS} \, \vartheta \, {}^{*}RR
    \right],
\end{align}
%
where $\vartheta$ is a pseudoscalar field which couples to the Pontryagin
density
${}^{*}RR = {}^{*}R_{\mu\nu}{}^{\rho\sigma} R^{\nu\mu}{}_{\rho\sigma}$,
%
where ${}^{*}R^{\mu\nu\rho\sigma}$ is the dual of the Riemann tensor
defined as
%
${}^{*}R^{\mu\nu\rho\sigma} =
\epsilon^{\mu\nu}{}_{\gamma\delta}
R^{\gamma\delta\rho\sigma} / 2$,
%
and $\epsilon_{\mu\nu\gamma\delta}$ is the Levi-Civita tensor.
%
The Pontryagin density by itself, $\int \dV {}^{*}RR$, is a boundary
term that does not contribute to the field equations in
four-dimensions~\cite{Jackiw:2003pm}.
%
The Pontryagin density can, however, modify the field equation when coupled to $\vartheta$;
the strength of the coupling set by $\ell_{\rm dCS}$, with dimensions of length.
%
% However, a coupling with $\vartheta$ can modify the field equations.
% %
% The coupling strength is set by the coupling constant $\ell_{\rm dCS}$, with
% dimensions of length.

@@

\begin{table*}[t]
% \begin{tabular}{c | c c c c c c c c c c}
% \hline
% \hline
% Theory      & $\omega^{(0)}_{220}$ & $\tau^{(0)}_{220}$ & $\omega^{(1)}_{220}$ & $\tau^{(1)}_{220}$ & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
% \hline
% GR          & 0.3737               & 11.2407            &  0.1258              & 0.2522             & --  & --                          & --                        & --                          & --                        & \cite{Maselli:2019mjd} \\
% \hline
% EdGB        & --                   & --                 &  --                  & --                 & 4   & 0.0107                      & 0.0044                    & $-$0.2480                   & $-$1.101                  & \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
% dCS         & --                   & --                 &  --                  & --                 & 4   & 3.1964                      & 6.3619                    & 41.199                      & 794.66                    & \cite{Wagle:2021tam}  \\
% cubic EFT   & --                   & --                 &  --                  & --                 & 4   & $-$0.5813                   & 2.6469                    & $-$3.8620                   & 265.12                    & \cite{Cano:2021myl} \\
% quartic EFT & --                   & --                 &  --                  & --                 & 6   & $-$0.2114                   & $-$0.6070                 & $-$1.5263                   & 171.35                    & \cite{Cano:2021myl}  \\
% \hline
% \hline
% \end{tabular}
%
\begin{tabular}{c | c }
\hline
\hline
                        & GR~\cite{Maselli:2019mjd} \\
\hline
$\omega^{(0)}_{220}$    & 0.3737                    \\
$\tau^{(0)}_{220}$      & 11.2407                   \\
$\omega^{(1)}_{220}$    & 0.1258                    \\
$\tau^{(1)}_{220}$      & 0.2522                    \\
\hline
\hline
\end{tabular}
\quad
\begin{tabular}{c | c  c  c  c}
\hline
\hline
                              & EdGB ($p=4$)~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} & dCS ($p=4$)~\cite{Wagle:2021tam} & cubic EFT ($p=4$)~\cite{Cano:2021myl} & quartic EFT ($p=6$)~\cite{Cano:2021myl} \\
\hline
$\delta\omega^{(0)}_{220}$    & 0.0107                                               & 3.1964                   & $-$0.5813                     & $-$0.2114                       \\
$\delta\tau^{(0)}_{220}$      & 0.0044                                               & 6.3619                   & $-$0.2114                     & $-$0.6070                       \\
$\delta\omega^{(1)}_{220}$    & $-$0.2480                                            & 41.199                   & $-$3.8620                     & $-$1.5263                       \\
$\delta\tau^{(1)}_{220}$      & $-$1.1010                                            & 794.66                   & 265.12                        & 171.35                          \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
%
We also included for comparison the GR coefficients, up to $j=1$, obtained in Ref.~\cite{Maselli:2019mjd}.
%
The remaining GR coefficients for $1< j \leqslant 4$, for which their non-GR counterpart cannot be determined as of yet for
the theories under consideration, can be found in Table~I of Ref.~\cite{Maselli:2019mjd}.
}
\label{tab:ref_theories_qnms}
\end{table*}

@@

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that it should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to each theory's fundamental length-scale $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
the Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final results using
$\varepsilon = 1/2$.
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %
We say that a bound has been placed on $\ell_{\rm th}$, if most of the PDF $P(\ell_{\rm th} | d)$ support is in the interval
$[0, \, \Lambda_{\rm EFT}(1/2, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell_{\rm th} | d)$, namely
%
\begin{equation}
P(\ell_{\rm th} \leqslant \ell^{\rm max}_{\rm th} | d) = \int_{0}^{\,\ell^{\rm max}_{\rm th}} \dd \ell'\, P(\ell' | d).
\label{eq:def_cdf}
\end{equation}
%
For instance, we require that for a bound at 90\% credible level to be placed on $\ell_{\rm th}$ that
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell_{\rm th} \leqslant \Lambda_{\rm EFT} | d) \leqslant 0.9,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
where we let $\ell^{\rm max}_{\rm th} = \Lambda_{\rm EFT}$ in Eq.~\eqref{eq:def_cdf}, and
likewise for other credibility percentiles.

@@

We start with EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the marginalized PDFs of the
coupling constant $\ell_{\rm GB}$, for GW150914 (top panel) and GW200129
(middle panel), with and without the spin corrections to the $(2,2,0)$ QNM.
%
The bottom panel shows the joint posterior obtained by combining both events.
%
We see, for both events, the $N_{\rm max} = 0$ posteriors are
characterized by a peak away from zero.
%
This does not mean that we are inferring a deviation from GR.
%
We recall that the deviations from GR in the ParSpec framework are controlled
by the dimensionless parameter $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm EdGB} = \left(\frac{c^2 \ell_{\rm EdGB}}{ G M^{\rm s}_{\rm f}}\right)^4\,.
    \label{eq:def_gamma_gb}
\end{equation}
%
As shown in Fig.~\ref{fig:gamma_gb}, $\gamma_{\rm EdGB}$ does indeed have a posterior
distribution with largest support at zero, indicating consistency between
the underlying signal and GR.
%
We also observe that the inclusion of the spin corrections (i.e., the curves with $N_{\rm max} = 1$)
displaces the posteriors distributions towards smaller values of $\ell_{\rm EdGB}$ (see Fig.~\ref{fig:edgb_exec_sum}),
and larger values of $\gamma_{\rm EdGB}$ (see Fig.~\ref{fig:gamma_gb}).

@@

As we have emphasized in Sec.~\ref{sec:remarks}, we must first check whether the
``EFT''~\eqref{eq:eft_bound} and ``ParSpec''~\eqref{eq:parspec_bound} bounds
are satisfied, before drawing any conclusions on the allowed values for
$\ell_{\rm EdGB}$ from our parameter estimations.
%
We check the validity of the EFT bound in Fig.~\ref{fig:edgb_cdf}. In the top (bottom) panel we show
the CDF of the $\ell_{\rm EdGB}$ posteriors for GW150914 (GW200129), obtained
by evaluating the integral~\eqref{eq:def_cdf} with $\ell^{\rm max}_{\rm th} = \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or
the remnant's mass ($\mathfrak{m} = M_{\rm f}$, solid lines), while varying $\varepsilon$ between 0 and 1.
%
For GW150914, we see that for the $N_{\rm max} = 0$ curves, that the CDF never goes
past $0.2$, regardless of the mass scale $\mathfrak{m}$ used and even at $\varepsilon =
1$, at which the EFT description of the theory would not be valid anyway.
%
This shows that that the ``EFT bound'' given by Eq.~\eqref{eq:eft_bound} is
never met to a significant credible level and thus that we cannot place a bound on
$\ell_{\rm EdGB}$.
%
The situation is similar for GW200129 with $N_{\rm max} = 0$ and does not
change for either event when we add spin corrections to the EdGB QNM.
%
For the case with $N_{\rm max} = 1$, we find that the ``EFT bound'' is satisfied
only for $\varepsilon \approx 0.8$ and $\approx1$ for GW150914 and GW200129, respectively.
However, we set the maximum value of $\varepsilon$ to be 1/2, thus,
taken together we are led to \emph{conclude that we cannot constrain EdGB gravity with our
present model}.
%
We summarize our findings in Table~\ref{tab:summary_edgb}.

@@

For dCS gravity, we follow Ref.~\cite{Wagle:2021tam}, which numerically calculated the QNMs
of slowly-rotating BHs, and found that for the axial gravitational-led modes the damping time increases, as we
increase $\ell_{\rm dCS}$, at constant BH spin.
%
Hence, according to the prescription of Sec.~\ref{sec:theory_specific_qnm},
this is the branch of QNMs we choose to work with.
